---
title: "TidyTuesdayGovtSpendingGit"
author: "Jeff Grayum"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

tuesdata <- tidytuesdayR::tt_load('2020-09-15')
kids <- tuesdata$kids
```

```{r}
library(tidyverse)
library(tidytuesdayR)
library(scales)
#install.packages("ggthemes")
library(ggthemes)
install.packages("janitor")
library(janitor)
theme_set(theme_light())
```

```{r}
kids %>%
  view()

kids %>%
  distinct(variable)

```

Some ideas!
-Principal component analysis: What are the dominant types of shape?
-Chloropleth (colored map by state).
-Animated map showing change from 1997-2016.
-What states had the fastest growing/shrinking in various variables?

We will start by considering only one variable, PK12ed.

```{r}
kids %>%
  filter(variable == "PK12ed") %>%
  group_by(year) %>%
  summarize(Total = 1000 * sum(inf_adj)) %>%
  ggplot(aes(year, Total)) +
  geom_line() +
  expand_limits(y = 0) +
  scale_y_continuous(labels = dollar)

kids %>%
  filter(variable == "PK12ed", 
         state %in% c("New York", "Florida", "Texas", "Rhode Island", "Washington")) %>%
  ggplot(aes(year, inf_adj_perchild, color = state)) +
  geom_line() +
  expand_limits(y = 0) 

#Now lets try the same thing, but sample 30 random states, and facet wrap.

kids %>%
  filter(variable == "PK12ed", 
         state %in% sample(unique(state), 30)) %>%
  mutate(state = fct_reorder(state, inf_adj_perchild, max, .desc = TRUE)) %>%
  ggplot(aes(year, 1000 * inf_adj_perchild)) +
  geom_line() +
  geom_vline(xintercept = 2009, color = "red", lty = 2) +
  scale_y_continuous(labels = dollar) +
  expand_limits(y = 0) +
  facet_wrap(~ state) +
  theme_tufte() +
  labs(x = "Year", y = "Inflation-adjusted Spending per Child") +
  theme(axis.text.x = element_text(angle = 90, hjust = -1))

#Now with all states.

kids %>%
  filter(variable == "PK12ed") %>%
  mutate(state = fct_reorder(state, inf_adj_perchild, max, .desc = TRUE)) %>%
  ggplot(aes(year, 1000 * inf_adj_perchild)) +
  geom_line() +
  geom_vline(xintercept = 2009, color = "red", lty = 2) +
  scale_y_continuous(labels = dollar) +
  expand_limits(y = 0) +
  facet_wrap(~ state) +
  theme_tufte() +
  labs(x = "Year", y = "Inflation-adjusted Spending per Child") +
  theme(axis.text.x = element_text(angle = 90, hjust = -1))

kids %>%
  filter(variable == "PK12ed") %>% 
  group_by(state) %>%
  mutate(change = inf_adj_perchild / first(inf_adj_perchild) -1) %>%
  ungroup() %>%
  mutate(state = fct_reorder(state, change, max, .desc = TRUE)) %>%
  arrange(state, year) %>% 
  ggplot(aes(year, change)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_line() +
  geom_vline(xintercept = 2009, color = "red", lty = 2) +
  scale_y_continuous(labels = percent, breaks = c(0, 1, 2)) +
  facet_wrap(~ state) +
  theme_tufte() +
  labs(x = "", y = "Increase in Inflation-adjusted Spending per Child Relative to 1997", title = "How has K-12 Spending Changed per State?", subtitle = "Red line represents global financial crash of 2009") +
  theme(axis.text.x = element_text(angle = 90, hjust = -1))

#Same thing, but showing change in spending on higher education.

kids %>%
  filter(variable == "highered") %>% 
  group_by(state) %>%
  mutate(change = inf_adj_perchild / first(inf_adj_perchild) -1) %>%
  ungroup() %>%
  mutate(state = fct_reorder(state, change, max, .desc = TRUE)) %>%
  arrange(state, year) %>% 
  ggplot(aes(year, change)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_line() +
  geom_vline(xintercept = 2009, color = "red", lty = 2) +
  scale_y_continuous(labels = percent, breaks = c(0, 1, 2)) +
  facet_wrap(~ state) +
  theme_tufte() +
  labs(x = "", y = "Increase in Inflation-adjusted Spending on Higher Education Relative to 1997", title = "How has Spending on Higher Education Changed per State?", subtitle = "Red line represents global financial crash of 2009") +
  theme(axis.text.x = element_text(angle = 90, hjust = -1))

#Showing changes in spending on social security, by state.

kids %>%
  filter(variable == "socsec") %>% 
  group_by(state) %>%
  mutate(change = inf_adj_perchild / first(inf_adj_perchild) -1) %>%
  ungroup() %>%
  mutate(state = fct_reorder(state, change, max, .desc = TRUE)) %>%
  arrange(state, year) %>% 
  ggplot(aes(year, change)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_line() +
  geom_vline(xintercept = 2009, color = "red", lty = 2) +
  scale_y_continuous(labels = percent, breaks = c(0, 1, 2)) +
  facet_wrap(~ state) +
  theme_tufte() +
  labs(x = "", y = "Increase in Social Security spending per Child Relative to 1997", title = "How has Social Security Spending per Child Changed by State?", subtitle = "Red line represents global financial crash of 2009") +
  theme(axis.text.x = element_text(angle = 90, hjust = -1))

kids %>%
  filter(variable == c("socsec", "PK12ed", "highered")) %>% 
  group_by(state) %>%
  mutate(change = inf_adj_perchild / first(inf_adj_perchild) -1) %>%
  ungroup() %>%
  mutate(state = fct_reorder(state, change, max, .desc = TRUE)) %>%
  arrange(state, year) %>% 
  ggplot(aes(year, change)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_line() +
  geom_vline(xintercept = 2009, color = "red", lty = 2) +
  scale_y_continuous(labels = percent, breaks = c(0, 1, 2)) +
  facet_wrap(~ state) +
  theme_tufte() +
  labs(x = "", y = "Increase in Social Security spending per Child Relative to 1997", title = "How has Social Security Spending per Child Changed by State?", subtitle = "Red line represents global financial crash of 2009") +
  theme(axis.text.x = element_text(angle = 90, hjust = -1))

#Lets write a function!

plot_change_faceted <- function(tbl){
tbl %>%
  group_by(state, variable) %>%
  mutate(change = inf_adj_perchild / first(inf_adj_perchild) -1) %>%
  ungroup() %>%  
  plot_faceted(change)+
  scale_y_continuous(labels = percent)  
}

plot_faceted <- function(tbl, y.axis){
  tbl %>%  
  mutate(state = fct_reorder(state, {{y.axis}}, change, max, .desc = TRUE)) %>%
  ggplot(aes(year, {{ y.axis }})) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_line() +
  facet_wrap(~ state) 
}

kids %>%
  filter(variable == "highered") %>% 
  plot_faceted(inf_adj_perchild) +
  geom_vline(xintercept = 2009, color = "red", lty = 2) +
  theme_tufte() +
  labs(x = "", 
       y = "Increase in inflation-adjusted spending per child on higher education relative to 1997", 
       title = "How has inflation-adjusted spending on higher education changed per state?", 
       subtitle = "Red line represents global financial crash of 2009") +
  theme(axis.text.x = element_text(angle = 90, hjust = -1))

```









```

